---
title: "Data Processing"
author: "Ben Gaiser & Jeremy Russell"
date: "10/7/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executing the Data Gathering File

```{r message=FALSE}
source('C:/Users/Benji/Desktop/Statistics/Git/Repositories/CSSR/DataGathering.R')
```


# Data Frame 1: Alcohol Consumption
## Units: Avergae serving sizes per person
## Source: 538 from World Health Organisation, Global Information System on Alcohol and Health (GISAH), 2010
## Variables of interest: 
### Independent Variable X: beer servings
### Dependent Variable Y: total litres of pure alcohol



## What the data look like - Initial Descriptive Statistics

```{r message=FALSE}
summary(AlcoholConsumption)
describe(AlcoholConsumption)
```

## Who drinks the most in total litres of pure alcohol?

```{r message=FALSE}
which.max(AlcoholConsumption$total_litres_of_pure_alcohol)
```

## The answer is row 16, i.e. Belarus

```{r message=FALSE}
head(AlcoholConsumption[16,])
```

## Who drinks the least in total litres of pure alcohol?

```{r message=FALSE}
which.min(AlcoholConsumption$total_litres_of_pure_alcohol) 
```

## The answer is row 1, i.e. Afghanistan
```{r message=FALSE}
head(AlcoholConsumption[1,])
```

## Hypothesis: Beer is the main driver of total litres of pure alcohol consumed
## per country

## Looking at the correlation to see if hypothesis is accurate

```{r message=FALSE}
cor(AlcoholConsumption$beer_servings, AlcoholConsumption$total_litres_of_pure_alcohol)    # 0.84
cor(AlcoholConsumption$wine_servings, AlcoholConsumption$total_litres_of_pure_alcohol)    # 0.67
cor(AlcoholConsumption$spirit_servings, AlcoholConsumption$total_litres_of_pure_alcohol)  # 0.65
```

## It seems that beer servings are most closely correlated to total litres of
## pure alcohol

## Plotting our findings in a scatterplot with a line of best fit and the 95%
## confidence interval

## Step 1: Creating a function for plotting a ggplot

```{r message=FALSE}
ggplotRegAlcCons <- function(fit){
  ggplot(AlcoholConsumption, aes(beer_servings, total_litres_of_pure_alcohol)) +
    geom_point(colour = 'blue') +
    stat_smooth(method = 'lm', col = 'red', size=0.75) +
    labs(title = paste('Adj R2 = ',signif(summary(fit)$adj.r.squared, 3),
                       'Intercept =',signif(fit$coef[[1]],3 ),
                       ' Slope =',signif(fit$coef[[2]], 1),
                       ' P =',signif(summary(fit)$coef[2,4], 2)))
}
```

## Step 2: Running the linear regression for the line of best fit

```{r message=FALSE}
FitOfData <- lm(total_litres_of_pure_alcohol ~ beer_servings, data=AlcoholConsumption)
```

## Step 3: Plotting the graph

```{r message=FALSE}
ggplotRegAlcCons(FitOfData)
```

## Analyzing five countries of interest to see how they differ in their beer servings values

## Finding Germany, USA, South Africa, China and Australia

```{r message=FALSE}
which(grepl('Germany', AlcoholConsumption$country)) # row 66
which(grepl('USA', AlcoholConsumption$country)) # row 185
which(grepl('South Africa', AlcoholConsumption$country)) # row 160
which(grepl('China', AlcoholConsumption$country)) # row 37
which(grepl('Australia', AlcoholConsumption$country)) # row 9
```

## Subsetting the Data 'AlcoholConsumption' for ease of commanding

```{r message=FALSE}
SubsetOfFiveCountries <- AlcoholConsumption[c(9, 37, 66, 160, 185),]
```

## Plotting our findings in a Scatterplot, we see that Germany drinks most out of the five countries on both variables

```{r message=FALSE}
ggplot(SubsetOfFiveCountries,
       aes(beer_servings, total_litres_of_pure_alcohol)) + 
  geom_point(aes(colour = factor(country))) +
  scale_colour_discrete(name='Countries')
```

# Data Frame 2: Swiss Data Set
## Source: Swiss Fertility and Socioeconomic Indicators (1888), R Data Set
## Variables of interest: 
### Independent Variable X: Catholic - % as opposed to Protestant
### Dependent Variable Y: Fertility - lg, common standardized fertility measure

## What the data look like - Initial Descriptive Statistics
summary(swiss)

## Hypothesis: Catholics have a higher fertility rate than Protestants

## A closer look at the initial descriptive statistics of our variables of interest

```{r message=FALSE}
describe(swiss$Fertility)
describe(swiss$Catholic)
var(swiss$Fertility)
var(swiss$Catholic)
sd(swiss$Fertility)
sd(swiss$Catholic)
```

## Catholic shows high variance and standard deviation for a continuous variable of between 0 and 100

## Plotting both variables to see their relationship

```{r message=FALSE}
ggplot(swiss, aes(Catholic, Fertility)) + geom_point()
```

## Plot fertility and Catholic with ggplot

```{r message=FALSE}
ggplotRegSwiss <- function(fit){
  ggplot(swiss, aes(Catholic, Fertility)) +
    geom_point(colour = 'blue') +
    stat_smooth(method = 'lm', col = 'red', size=0.75) +
    labs(title = paste('Adj R2 = ',signif(summary(fit)$adj.r.squared, 3),
                       'Intercept =',signif(fit$coef[[1]],3 ),
                       ' Slope =',signif(fit$coef[[2]], 1),
                       ' P =',signif(summary(fit)$coef[2,4], 2)))
}
FitOfDataSwiss <- lm(Fertility ~ Catholic, data = swiss)
ggplotRegSwiss(FitOfDataSwiss)
```

## Despite outliers, this is still a significant relationship, explaining 20 % of the variance (R-value)

## Which cantons are neither mostly Protestant nor mostly Catholic?

## Step 1: Creating a factor variable with four different groups

```{r message=FALSE}
swiss$CatholicCat <- cut(swiss$Catholic, seq(0, 100, 25))
```

## Step 2: Changing the factor variable into a character variable for renaming the rownames

```{r message=FALSE}
swiss$CatholicCat <- as.character(swiss$CatholicCat)
```

## Step 3: Renaming the rownames

```{r message=FALSE}
swiss$CatholicCat[swiss$CatholicCat=='(0,25]'] <- 'Protestant'
swiss$CatholicCat[swiss$CatholicCat=='(25,50]'] <- 'Protestant to Catholic'
swiss$CatholicCat[swiss$CatholicCat=='(50,75]'] <- 'Catholic to Protestant'
swiss$CatholicCat[swiss$CatholicCat=='(75,100]'] <- 'Catholic'
```

## Step 4: Finding the cantons which are 'Protestant to Catholic' and 'Catholic to Protestant'

```{r message=FALSE}
which(grepl('Protestant to Catholic', swiss$CatholicCat)) # 4 and 45
which(grepl('Catholic to Protestant', swiss$CatholicCat)) # 46 and 47
```

## Step 5: Searching the names

```{r message=FALSE}
swiss[c(4,45:47),]
```

## Moutier, V. De Geneve, Rive Droite and Rive Gauce are the only cantons where there is at least a third of the population that does not belong to the majority religion




